---
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
    - \usepackage{fancyhdr}
    - \usepackage{fontspec}
    - \usepackage{xcolor}
    - \geometry{left = 0.5in,right = 0.5in}
    - \usepackage{floatrow}
    - \usepackage{caption}
#mainfont: Roboto
#sansfont: Roboto
urlcolor: purplepeopleeater
---

```{r, echo=FALSE, include=FALSE}

library(crayon)
library(tidyverse)

#Importing Data Table

#runresults <- read.csv("sequencerun_data.csv")
runresults <- read.csv("Sample_table_2021-11-01.csv")

runresults$batch_id <- str_remove(runresults$batch_id, "batch_")

```

<!-- define color and adjust lengths for header and footer-->
\definecolor{purplepeopleeater}{RGB}{106,13,75}
\addtolength{\headheight}{3.0cm}
\addtolength{\topmargin}{-0.5in}
\addtolength{\footskip}{-0.225in}

<!-- % setup header -->
\pagestyle{fancy}
\fancyhf{}

<!-- header content -->
<!-- Uncomment the line of code below to include a header -->
<!-- \fancyhead[L]{\raisebox{-0.05\height}{\includegraphics[height = 2.5cm]{"Commonwealth_MA_sign.png"}}}-->
\fancyhead[L]{\Huge MASPHL Genomic Sequencing Run Report\\
\Large `r paste("Report Date:", Sys.Date(), "for batch", tail(runresults$batch_id, n=1), "sequenced", tail(runresults$seq_date, n=1)) `}

<!-- create red header line -->
\renewcommand{\headrulewidth}{1pt}
\renewcommand{\headrule}{\hbox to\headwidth{%
    \color{purplepeopleeater}\leaders\hrule height \headrulewidth\hfill}}

<!-- footer content -->
\fancyfoot[C]{For research use only, not for clinical use.}
\fancyfoot[R]{\thepage}

<!-- create red footer line -->
\renewcommand{\footrulewidth}{1pt}
\renewcommand{\footrule}{\hbox to\headwidth{%
    \color{purplepeopleeater}\leaders\hrule height \headrulewidth\hfill}}


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

library(lubridate)
library(dplyr)
library(data.table)
library(scales)
library(tibble)
library(tidyverse)


#converting the dates from character to numeric
runresults$collection_date <- as_date(runresults$collection_date, format = "%m/%d/%Y")
runresults$seq_date <- as_date(runresults$seq_date, format = "%m/%d/%Y")

#Flagging for any results that are out of the date range
daterange <- function(t) {
  ifelse(t < 2020-01-01, 1, 0)}


#code excludes the sequences that were not run (collection date = blank)
runresults2 <- runresults %>% filter(!(is.na(collection_date)))  %>%
      mutate(Assembly_Count = case_when(assembly_status %in% "PASS" ~ 1, assembly_status %in% "FAIL" ~ 0)) %>%
  mutate(Assembly_Count_Fail = case_when(assembly_status %in% "FAIL" ~ 1, assembly_status %in% "PASS" ~ 0)) %>%
      mutate(percent_SC2_reads = case_when(X._SARS.COV.2_reads == 'TBD' ~ NA_real_, TRUE ~ as.numeric(X._SARS.COV.2_reads))) %>%
      mutate(percent_human_reads = case_when(X._human_reads == 'TBD' ~ NA_real_, TRUE ~ as.numeric(X._human_reads)))


#creating a summary table

batchsummary <- runresults2 %>% group_by(batch_id) %>% summarise(n = n(), MinDate = min(collection_date, na.rm = TRUE), MaxDate = max(collection_date, na.rm = TRUE), pass_number = sum(Assembly_Count), fail_number = sum(Assembly_Count_Fail))

#Getting the median and IQR for the pass rate for later analysis
summary(batchsummary$Pass_Rate)

batchsummary <- batchsummary %>% mutate(Pass_Rate = pass_number/n*100) %>% mutate(Pass_Group = case_when(Pass_Rate <= quantile(Pass_Rate, 0.25) ~ "Q1", quantile(Pass_Rate, 0.25) < Pass_Rate & Pass_Rate <= median(Pass_Rate) ~ "Q2",median(Pass_Rate) < Pass_Rate & Pass_Rate <= quantile(Pass_Rate, 0.75) ~ "Q3",quantile(Pass_Rate, 0.75) < Pass_Rate ~ "Q4"))

a <- batchsummary %>% select(batch_id, Pass_Group)

runresults_join <- left_join(runresults2, a, by = "batch_id") %>% mutate(human_reads = (percent_human_reads * fastqc_raw_reads_1)/100) %>% mutate(SC2_reads = (percent_SC2_reads * fastqc_raw_reads_1)/100)


rr_clean <- runresults_join %>% select(batch_id, seq_date, sample_id, collection_date, assembly_status, Assembly_Count, Assembly_Count_Fail, pangolin_lineage, nextclade_lineage, fastqc_raw_reads_1, pairs_kept_after_cleaning, percent_kept_after_cleaning, coverage_after_trimming, num_ACTG, num_N, meanbaseq_trim, meanmapq_trim, percent_human_reads, percent_SC2_reads) %>%  mutate(human_reads = (percent_human_reads * fastqc_raw_reads_1)/100) %>% mutate(SC2_reads = (percent_SC2_reads * fastqc_raw_reads_1)/100)


clean_batchsummary <- rr_clean %>% group_by(batch_id) %>% summarise(n = n(), Seqdate = max(seq_date),
                                                                MinDate = min(collection_date, na.rm = TRUE),
                                                                MaxDate = max(collection_date, na.rm = TRUE),
                                                                pass_number = sum(Assembly_Count),
                                                                fail_number = sum(Assembly_Count_Fail),
                                                                Median_rawread = median(fastqc_raw_reads_1, na.rm = TRUE),
                                                                rawreadQ25 = quantile(fastqc_raw_reads_1, probs = 0.25, na.rm = TRUE),
                                                                rawreadQ75 = quantile(fastqc_raw_reads_1, probs = 0.75, na.rm = TRUE),

                                                                 Median_cleanread = median(pairs_kept_after_cleaning, na.rm = TRUE), 
                                                                cleaningQ25 = quantile(pairs_kept_after_cleaning, probs = 0.25, na.rm = TRUE),
                                                                cleaningQ75 = quantile(pairs_kept_after_cleaning, probs = 0.75, na.rm = TRUE),

                                                                 Median_trim = median(coverage_after_trimming, na.rm = TRUE),
                                                                coveragetrimQ25 = quantile(coverage_after_trimming, 0.25, na.rm = TRUE),
                                                                 coveragetrimQ75 = quantile(coverage_after_trimming, 0.75, na.rm = TRUE),

                                                                Median_ACTG = median(num_ACTG, na.rm = TRUE),
                                                                ACTG_Q25 = quantile(num_ACTG, probs = .25, na.rm = TRUE),
                                                                  ACTG_Q75 = quantile(num_ACTG, probs = .75, na.rm = TRUE),

                                                                Median_N = median(num_N, na.rm = TRUE),
                                                                N_Q25 = quantile(num_N, probs = 0.25, na.rm = TRUE),
                                                                 N_Q75 = quantile(num_N, probs = 0.75, na.rm = TRUE),

                                                                Median_humanreads = median(human_reads, na.rm = TRUE),
                                                                humanreasd_Q25 = quantile(human_reads, probs = .25, na.rm = TRUE),
                                                                humanreasd_Q75 = quantile(human_reads, probs = .75, na.rm = TRUE),

                                                                Median_SC2reads = median(SC2_reads, na.rm = TRUE),
                                                                SC2reads_Q25 = quantile(SC2_reads, probs = 0.25, na.rm = TRUE),
                                                                SC2reads_Q75 = quantile(SC2_reads, probs = 0.75, na.rm = TRUE))

clean_batchsummary_table <- clean_batchsummary %>% mutate(SampleRange = paste(MinDate, MaxDate, sep = " - ")) %>% mutate(Per_Passing = round((pass_number/n*100), digits = 2))

clean_batchsummary_table <- as.data.frame(clean_batchsummary_table)
clean_batchsummary_table <- setorder(setDT(clean_batchsummary_table), cols = Seqdate)[]
clean_batchsummary_table <- rownames_to_column(clean_batchsummary_table, "Batch")
clean_batchsummary_table$Batch <- as.integer(clean_batchsummary_table$Batch)

#total
clean_batchsummary_total <- rr_clean %>% summarise(n = n(), Seqdate = max(seq_date),
                                                                MinDate = min(collection_date, na.rm = TRUE),
                                                                MaxDate = max(collection_date, na.rm = TRUE),
                                                                pass_number = sum(Assembly_Count),
                                                                Median_rawread = median(fastqc_raw_reads_1, na.rm = TRUE),
                                                                rawreadQ25 = quantile(fastqc_raw_reads_1, probs = 0.25, na.rm = TRUE),
                                                                rawreadQ75 = quantile(fastqc_raw_reads_1, probs = 0.75, na.rm = TRUE),

                                                                 Median_cleanread = median(pairs_kept_after_cleaning, na.rm = TRUE),
                                                                cleaningQ25 = quantile(pairs_kept_after_cleaning, probs = 0.25, na.rm = TRUE),
                                                                cleaningQ75 = quantile(pairs_kept_after_cleaning, probs = 0.75, na.rm = TRUE),

                                                                 Median_trim = median(coverage_after_trimming, na.rm = TRUE),
                                                                coveragetrimQ25 = quantile(coverage_after_trimming, 0.25, na.rm = TRUE),
                                                                 coveragetrimQ75 = quantile(coverage_after_trimming, 0.75, na.rm = TRUE),

                                                                Median_ACTG = median(num_ACTG, na.rm = TRUE),
                                                                ACTG_Q25 = quantile(num_ACTG, probs = .25, na.rm = TRUE),
                                                                  ACTG_Q75 = quantile(num_ACTG, probs = .75, na.rm = TRUE),

                                                                Median_N = median(num_N, na.rm = TRUE),
                                                                N_Q25 = quantile(num_N, probs = 0.25, na.rm = TRUE),
                                                                 N_Q75 = quantile(num_N, probs = 0.75, na.rm = TRUE),

                                                                Median_humanreads = median(human_reads, na.rm = TRUE),
                                                                humanreasd_Q25 = quantile(human_reads, probs = .25, na.rm = TRUE),
                                                                humanreasd_Q75 = quantile(human_reads, probs = .75, na.rm = TRUE),

                                                                Median_SC2reads = median(SC2_reads, na.rm = TRUE),
                                                                SC2reads_Q25 = quantile(SC2_reads, probs = 0.25, na.rm = TRUE),
                                                                SC2reads_Q75 = quantile(SC2_reads, probs = 0.75, na.rm = TRUE))
clean_batchsummary_total <- clean_batchsummary_total %>% mutate(SampleRange = paste(MinDate, MaxDate, sep = " - ")) %>% mutate(Per_Passing = pass_number/n*100)
clean_batchsummary_total <- as.data.frame(clean_batchsummary_total)
clean_batchsummary_total <- setorder(setDT(clean_batchsummary_total), cols = Seqdate)[]
clean_batchsummary_total <- rownames_to_column(clean_batchsummary_total, "Batch")
clean_batchsummary_total$Batch <- 0


clean_batchsum_all <- rbind(clean_batchsummary_table, clean_batchsummary_total, fill=TRUE)

lastfive <- rr_clean %>% filter(batch_id %in% tail(clean_batchsummary_table$batch_id,n=8))

Assemblies <- rr_clean %>% group_by(seq_date, batch_id,assembly_status) %>% summarise(n = n())
Assemblies <- rownames_to_column(Assemblies, "Row")
x <- rr_clean %>% group_by(batch_id) %>% summarise(n_batch =n())
Assemblies <- left_join(Assemblies, x, by = "batch_id")
Assemblies$Percent <- round((Assemblies$n/Assemblies$n_batch)*100, digits = 2)

```

# Current Run Statistics
```{r, include=FALSE}

Current_Run <- runresults %>% filter(seq_date==max(seq_date)) %>%
  mutate(Limit = case_when(sample_id %like% "HSC" ~ "100",
                           sample_id %like% "NEG" ~ "100",
                           sample_id %like% "NTC" ~ "600",
                           TRUE ~ ""))
Current_Run$Limit <- as.integer(Current_Run$Limit)
Current_Run$Control_Status <- if_else(Current_Run$Limit <= Current_Run$num_ACTG, "Fail","Pass")


```

\floatsetup[figure]{style=plaintop}
\newfloatcommand{btabbox}{table}

\begin{figure}\CenterFloatBoxes
  \begin{floatrow}
  \captionsetup{font={Large}}
    \ffigbox{%

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="100%"}

g <- ggplot(data = Current_Run, aes(x=collection_date)) + geom_histogram(aes(x = collection_date), binwidth=1) +
              scale_x_date(labels = date_format("%b %d, %y"), limits = c(min(Current_Run$collection_date), Current_Run$seq_date), date_breaks = "1 week")  +
                    ylab("Number of Isolates") + xlab("Collection Date of Isolate") +
                          geom_vline(aes(xintercept = seq_date), linetype = "dashed") +
                               geom_text(aes(x=seq_date, y = 0,label="Sequencing Date", family = "serif"), size = 6, angle = 90, vjust = -0.4,hjust = -0.75) +
                                            theme_classic()  +
                                                  theme(text = element_text(family="serif"),plot.title = element_text(hjust=0.5) ,axis.title.x = element_text(size = 24), axis.title.y = element_text(size=24), axis.text=element_text(size=16), axis.text.x = element_text(angle = 45, hjust=1))

g

#g + geom_vline(aes(xintercept = seq_date), linetype = "dashed") +
#                                      geom_text(aes(x=seq_date, y = 0,label="Sequencing Date", family = "serif"), size = 4.5, angle = 90, vjust = -0.4,hjust = -0.75)

```
    }{\caption*{Isolate Collection Dates}}
    \ttabbox{%
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='right'}

library(formattable)

Current_Run_Controls <- Current_Run %>% filter(seq_date==max(seq_date)) %>% filter(is.na(collection_date)) %>% select(sample_id, num_ACTG, Limit, Control_Status) %>% filter(!(str_detect(sample_id,"HSC")))

knitr::kable(Current_Run_Controls, "latex", col.names = c("Control ID", "ACTG Detected", "Limit", "Interpretation"), align = "c")

```
    }{\caption*{Contamination Stats}}
  \end{floatrow}
\end{figure}

# Current and Historic Trends
```{r, echo=FALSE, out.width="90%", warning=FALSE,message=FALSE}

cols <- c("gray78","blue")
names(cols) <- c("FAIL", "PASS")

#Creates a list of Unique batch IDs to controls for batches that have no "Pass" or no "Fail" sequences
Assemblies_Set <- Assemblies$batch_id %>% unique() %>% tail(8)

ggplot(Assemblies[which(Assemblies$batch_id %in% Assemblies_Set),], aes(x=Percent, y=batch_id, fill=assembly_status, color=batch_id)) +
  geom_bar(position="stack", stat="identity", size = 2) +
  scale_fill_manual(values=c(cols)) +
  scale_color_manual(values=c(rep("white", times=7),"black")) +
  ggtitle("Assemblies Status of Sequences") +
  labs( x= "Percent of Sequences in Batch", y = "Batch ID", caption = "Values in bars represent number of sequences in category") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size=16), axis.text.y=element_text(color=c(rep("black",times=10),"blue"),face="bold"), plot.caption = element_text(hjust = 0, size = 14),text = element_text(family="serif"), axis.title.y = element_text(vjust = 3, size=15),axis.title.x = element_text(size = 16)) +
  guides(fill=guide_legend(title="Assembly Status"), color = FALSE) +
  geom_text(aes(label=n),color="white", position = "stack", hjust= 1.4, face="bold")

#plot <- plot_ly(data=tail(clean_batchsummary_table,n=5), x = ~batch_id, y = ~pass_number, type = 'bar', name = 'Pass') %>%  add_trace(y = ~fail_number, name = 'Fail') %>%
    #layout(yaxis = list(title = 'Count'), barmode = 'stack')
```
\newpage

```{r,echo=FALSE, include=FALSE, message=FALSE, error=FALSE}

rawplot <- ggplot(lastfive, aes(x=batch_id, y=fastqc_raw_reads_1)) +
  geom_boxplot(aes(color=batch_id)) +
  scale_color_manual(values=c("black", "black", "black", "black", "black","black","black", "blue"))+
  labs(title = "Median Number of Raw Reads", x = "Batch ID", y = "Number of Reads") +
  theme_classic() +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(size = 10, face = "bold", angle=45), axis.title.x = element_text(size = 10), axis.title.y = element_text(size = 10), text = element_text(family="serif"))

clean_plot <- ggplot(lastfive, aes(x=batch_id, y=pairs_kept_after_cleaning)) +
  geom_boxplot(aes(color=batch_id)) +
  scale_color_manual(values=c("black", "black", "black", "black", "black","black","black", "blue"))+
  labs(title = "Median Number of Cleaned Read Pairs", x = "Batch ID", y = "Number of Read Pairs") +
  scale_y_continuous(labels = scales::comma) +
  theme_classic() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(size = 10, face = "bold",angle=45), axis.title.x = element_text(size = 10), axis.title.y = element_text(size = 10), text = element_text(family="serif"))

```

```{r, echo=FALSE, include=FALSE, message=FALSE, error=FALSE}

library(tibble)


is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}

dat <- lastfive %>% tibble::rownames_to_column(var="outlier") %>% group_by(batch_id) %>% mutate(is_outlier=ifelse(is_outlier(human_reads), human_reads, as.numeric(NA)))
dat$sample_id[which(is.na(dat$is_outlier))] <- as.numeric(NA)

ggplot(dat, aes(y=human_reads, x=factor(batch_id))) + geom_boxplot() + geom_text(aes(label=sample_id),na.rm=TRUE,nudge_y=0.05)


human_plot <- ggplot(dat, aes(x=batch_id, y=human_reads)) +
  geom_boxplot(aes(color=batch_id)) +
   #geom_text(aes(label=sample_id),na.rm=TRUE,nudge_y=0.05) +
  scale_color_manual(values=c("black", "black", "black", "black", "black","black","black", "blue")) +
  labs(title = "Median Number of Human Reads", x = "Batch ID", y = "Number of Reads") +
  scale_y_continuous(labels = scales::comma, limits = c(0,1000000)) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(family="serif"), plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(size = 10, face = "bold",angle = 45), axis.title.x = element_text(size = 10), axis.title.y = element_text(size = 10))

dat2 <- lastfive %>% tibble::rownames_to_column(var="outlier") %>% group_by(batch_id) %>% mutate(is_outlier=ifelse(is_outlier(SC2_reads), SC2_reads, as.numeric(NA)))
dat$sample_id[which(is.na(dat$is_outlier))] <- as.numeric(NA)

SCOV_plot <- ggplot(dat2, aes(x=batch_id, y=SC2_reads)) +
  geom_boxplot(aes(color=batch_id)) +
  #geom_text(aes(label=sample_id),na.rm=TRUE,nudge_y=0.05) +
  scale_color_manual(values=c("black", "black", "black", "black", "black","black","black", "blue"))+
  labs(title = "Median Number of SARS-COV-2 Reads", x = "Batch ID", y = "Number of Reads") +
  scale_y_continuous(labels = scales::comma) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(family="serif"),plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(size = 10, face = "bold", angle=45), axis.title.x = element_text(size = 10), axis.title.y = element_text(size = 10))

library(cowplot)

```

```{r echo=FALSE, warning=FALSE, error=FALSE, fig.show="hold", fig.align='left', fig.width=8, fig.height=7}

#par(mar = c(4, 4, .1, .1))
#rawplot
#clean_plot

plot_grid(rawplot, clean_plot, human_plot, SCOV_plot, ncol = 2, nrow = 2, scale = 0.9)

```
