---
title: "Wastewater Variants Visualization"
author: Emily Tyszka & Brooke Talbot
output: html_document
date: '2023-12-20'
---

# Libraries
```{r}

library(stringr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly) ##make sure htmltools packages is up to date 


```


# Data
```{r cleaning, include=FALSE, echo=FALSE , warning=FALSE, message=FALSE, results = "asis", type = 'latex'}

# Load in Freyja data (a demixed and aggregated file)
input <-("C:/Users/Brooke/OneDrive/APHL_Genomics_Specialist/APHL_Genomics_Specialist/WWS/demixed_aggregate_12_23.tsv") 

#read as a table
data <-read.table(input, header = TRUE, sep = "\t")

acc <- data.frame(data$X)
names(acc)[1] <- "Accession Number"
acc$rn <- row.names(acc) #create a row name for accession numbers df


##Counting the maximum number of lineages in the data set 

leng_data <- c()
for (i in 1:nrow(data)) {

leng_data <- append(leng_data, print(length(gregexpr(" ", data$lineages)[[i]])))

}
n <- max(leng_data) + 1


#clean lineages data
Lineages <- str_split_fixed(data$lineages, " ", n) #split into new column every time there's a space (after a new listed lineage name). n is calculated from the input dataset to account for different strains potentially detected at different sampling effors

Lineages <- data.frame(Lineages)
Lineages$rn <- row.names(Lineages) #add row numbers
Lineages <- Lineages %>% left_join(acc, by = c("rn" = "rn")) #join accessions to lineages by row number
## reordering the index columns
Lineages <- Lineages %>% select(rn, `Accession Number`, X1:ncol(Lineages))

##Identifying the last column in the data
lastcol <- ncol(Lineages)

Lineages_long <- gather(Lineages, index, Lineage, X1:lastcol, factor_key=TRUE)  #convert long to short - X1 X2 etc labelled as "index". corresponds with the number of columns
Lineages_long$index2 <- paste(Lineages_long$`Accession Number`, Lineages_long$index, sep=".") #create secondary index (WW22005.X1, WW22005.X2 etc) - this is identical between freq and lineage tables and will be used to merge them
view(Lineages_long)

#clean lineage frequencies data
freqs <- str_split_fixed(data$abundances, " ", n) #all here is the same as above, just using the frequencies that pertain to the lineages
freqs <- data.frame(freqs)
freqs$rn <- row.names(freqs)
freqs <- freqs %>% left_join(acc, by = c("rn" = "rn"))
freqs <- freqs %>% select(rn, `Accession Number`, X1:ncol(freqs))

freqs_long <- gather(freqs, index, Frequency, X1:lastcol, factor_key=TRUE) #
freqs_long$index2 <- paste(freqs_long$`Accession Number`, freqs_long$index, sep=".")
freqs_long <- freqs_long %>% select(index2, Frequency)
view(freqs_long)

#merge together lineages and their frequencies
all <- Lineages_long %>% full_join(freqs_long, by = c("index2" = "index2")) #merge by the secondary index
all <- subset(all, select = - c(index, index2)) #cleaning by dropping merge-associated columns
all <- all %>% drop_na(Frequency) #drop NAs
all <- all[!is.na(as.numeric(all$Frequency)), ] 
all$Frequency <- as.numeric(all$Frequency)

all <- all %>% drop_na(Frequency) #drop NAs
view(all) 
```

## Variants of Concern, Variants of Interest, and Variants being monitored ##
```{r, include=FALSE}

## Creating a list of the variants of concern from CDC ##
## voc, VBM, and VOI as of 1 Sept 2023 ##

voc = c("BA.2.86",
        "XBB.1.9.1",
        "XBB.1.9.2",
        "XBB.2.3",
        "XBB.1.16",
        "XBB.1.16.6",
        "XBB.1.5",
        "CH.1.1",
        "BA.2.74",
        'B.1.1.7',
        'B.1.351',
        'P.1',
        'B.1.617.2',
        "B.1.427",
        "B.1.429",
        "B.1.525",
        "B.1.526",
        'B.1.617.1',
        'B.1.1.529',
        'P.2',
        'B.1.621',
        'B.1.621.1', 
        'EG.5',
        'FL.1.5.1',
        'BA.1')

## Another strategy is to keep a running list of variants of concern (historical and current), then use a subsetter, such as the name, type, or year it was selected as a filter. 


```

```{r cleaner Lineage plot, echo=FALSE , warning=FALSE, message=FALSE, results = "asis", type = 'latex', include=FALSE}

## Making an interactive plot since there are so many lineages

all$LineageClean <- substr(all$Lineage, 1, 4)
#all$LineageClean <- substr(all$Lineage, 1, 6)

## Adds a label 
all <- all %>%
  mutate(mylabel = paste("Lineage:", LineageClean))

p <- ggplot(all, aes(fill=LineageClean, y= Frequency, x=`Accession Number`, text = mylabel)) + 
  geom_bar(position='stack', stat='identity', show.legend = FALSE) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + theme(legend.key.size = unit(.4, 'cm'))+
  scale_y_continuous(limits = c(0, 1)) + labs(color = "Variants", title = "Frequency of Main Lineages Per Sample") 
```
```{r}
ggplotly(p, tooltip = c("mylabel"))

```

```{r cleaner Lineage plot at threshold, echo=FALSE , warning=FALSE, message=FALSE, results = "asis", type = 'latex'}


## Interactive Facet Plot, but still a little clunky
## filter to frequencies approaching or greater than 1 %
p2<- all %>% filter(Frequency >= 0.01) %>%
ggplot( aes(fill=Lineage, y= Frequency, x=`Accession Number`)) + 
  geom_bar(position='stack', stat='identity') + theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 1)) + labs(color = "Variants", title = "Sub Lineage Frequency >0.9%, By Major Lineage") + 
facet_wrap(~LineageClean) 
#ggsave("Facet_lineages_2023.png")

ggplotly(p2)


```

## Looking for sublineages that are averaging a higher percentage

```{r}
## First pass, see if there are any sublianges within all of the major lineages that have a higher frequency that 1% in each sequence

library(dplyr)
sumtable <- all%>% dplyr::group_by(LineageClean) %>% dplyr::summarise(count = n(), sublineage_mean = 100*mean(Frequency), Upper_SE = 100*((mean(Frequency) + sd(Frequency)/sqrt(length((Frequency))))), Lower_SE = 100*((mean(Frequency) - sd(Frequency)/sqrt(length((Frequency))))), maxium = 100*max(Frequency))

sumtable <- sumtable[order(-sumtable$sublineage_mean),]

sumtable

knitr::kable(sumtable, "html", col.names = c("Clean Lineage", "No. Samples", "Mean Abundance (%)", "Mean Upper SE ", "Mean Lower SE", "Maximum Abundance"), align = "c")



```

## Let's answer the question: Where do the variants of concern average in terms of dominance compared to other lineages? ##
```{r}

## Filtered for just the variants of Concern ##
all %>% filter(Lineage %in% voc) %>% ggplot(aes(fill=Lineage, y= Frequency, x=`Accession Number`)) + 
  geom_bar(position='stack', stat='identity') + theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 1)) + labs(color = "Variants", title = "Variants of Interest, Concern, and Monitored  Frequency Plot") + 
facet_wrap(~Lineage) 

## Looking at the emergence of Variants of Concern by time
#merge with dates for each accession
dates1<-("C:/Users/Brooke/OneDrive/APHL_Genomics_Specialist/APHL_Genomics_Specialist/WWS/wws_dates_12_23.tsv") #matches dates to accessions
dates<-read.table(dates1, sep = "\t", header = TRUE)
all_dates <- all %>% dplyr::full_join(dates, by = c("Accession Number" = "entity.wws_id")) #link data with date decoder ring

all_dates <- all_dates %>% mutate(label = paste("Collection Date:", collection_date))

pdates <- all_dates %>% filter(Lineage %in% voc) %>% ggplot(aes(fill=Lineage, y= Frequency, x=collection_date, text = label)) + 
  geom_bar(position='stack', stat='identity') + theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 1)) + labs(color = "Variants", title = "Variants of Interest, Concern, and Monitored  Frequency Plot by Date", x = "Collection Date") + 
facet_wrap(~Lineage) 

ggplotly(pdates, tooltip = c("label"))
